---
title: 'Tidy Forecasting in R: ETS & ARIMA Modeling'
author: "Joon Im"
date: 2020-10-18
categories: ["R","Forecasting","Time Series"]
tags: ["R", "tidyverse", "time Series", "modeltime", "fpp2", "fpp3", "forecast","fable"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
library(formatR)
library(tidyquant)
```

# 1. Set Up

### 1.1 Introduction

This article is the second in a series comparing the `fpp2`, `fpp3`, and `modeltime` forecasting frameworks in `R`. Specifically we'll be focusing on how to automatically choose both exponential smoothing and ARIMA models. 

### 1.2 Load Libraries

```{r libraries, message=FALSE, warning=FALSE}
# Load libraries
library(fpp2)         # The forecasting OG
library(fpp3)         # Official update to fpp2
library(modeltime)    # The forecasting newcomer
library(timetk)       # Companion to modeltime
library(parsnip)      # Common interface for specifying models
library(tidyverse)    # Data manipulation tools
library(cowplot)      # Arranging plots
```

### 1.3 About Exponentially Weighted Models

- A family of models where the **trend** component is either *missing*, *additive* or *damped* and the **seasonal** component is either *missing*, *additive*, or *multiplicative*
- Forecasts are weighted averages of past observations i.e. the more recent an observation, the higher the associated weight

> ETS model forecasts can be quickly generated and generalize well to many different types of time series

### 1.4 About AutoRegressive Integrated Moving Average Models

- A family of models where:
  - **AR**: variable of interest undergoes a regression against prior values
  - **I**: variable of interest undergoes a normalizing transformation called *differencing* for better model fitting
  - **MA**: variable of interest undergoes a regression against prior residual errors 
- Forecasts are based on the idea that models can find patterns from past values which are then applied to future values

# 2. An Overview of Exponential Smoothing Models

The following examples will utilize a dataset containing yearly winning times of the Boston marathon.

```{r}
# Load Boston marathon winning times 
m_tbl <- fpp2::marathon %>%
  tk_tbl(rename_index = "year") %>%
  mutate(year = ymd(year, truncated = 2L))

# Check it out
m_tbl %>% head()
```

### 2.1 Simple Exponential Smoothing

- The forecasts are simply the last value of the estimated level
- The level changes over time as a function of the most recent observation

```{r}
# Convert to tsibble
m_tsbl <- fpp2::marathon %>% as_tsibble()

# Fit a SES model: additive errors, no trend, no seasonality
fit <- m_tsbl %>%
  model(ets = ETS(value ~ error("A") + trend("N") + season("N")))

# Forecast for 10 years into the future
fc <- fit %>% forecast(h = 10)

# Plot forecasts + one-step forecasts for the training data
fc %>%
  autoplot(m_tsbl) +   # Plot actual data 
  geom_line(aes(y = .fitted, colour = "Fitted"),  
            data = augment(fit)) +  # Show how model fits the data
  labs(title="Boston marathon winning times with 10-year SES forecast",
       x="", y="Minutes") 
```

> 💡 If your data has no trend and no seasonality, then an SES model may work well.

### 2.2 Holt's Linear Method

- Holt’s linear method extends SES by adding trend
- The forecast equation is a linear function of the forecast horizon where forecasts are the last value of the estimated level
- The level also changes over time as a function of the most recent observation with added trend 






